---
title: "Big Picture"
description: "Mental Wellbeing, Access to Care and Factors Impacting the US Mental Health Crisis"
toc: true
draft: FALSE
filters:
  - shinylive
---

There are several different factors that impact both mental wellbeing and access to mental health care. Through analysis of the Household Pulse Survey and Area Health Resource Data, this website examines the relationship between mental well being, access to mental health care, age, income, education, and race. 

Mental well being is measured through the Household Pulse Survey where families reported the frequency of certain mental health symptoms such as loneliness, anxiety and feeling down. These were then totaled to find an overall measure of mental well being. Age, income and education level also come from the Household Pulse Survey. Access to mental health care is measured through the number of psychologists per 100k people. 


```{r}
#| echo: false
#| message: false
#| warning: false
#| results: 'hide'
#| column: screen
#| out-width: 100%

suppressMessages(suppressWarnings(library(tidyverse)))
suppressMessages(suppressWarnings(library(sf)))
suppressMessages(suppressWarnings(library(tidycensus)))

#Combining joint data with census data to create maps

joint_data <- read.csv("dataset/joint_data.csv")

joint_data <- joint_data |>
  group_by(region) |>
  reframe(mean_edu = mean(Educational_Attainment),
           children_need = sum(Children_Need_for_Mental_Health_Treatment, na.rm = TRUE),
           children_receive = sum(Children_Receive_Mental_Health_Treatment, na.rm = TRUE),
           mean_income = mean(Household_Income, na.rm = TRUE),
           mean_total_mh = mean(total_mentalhealth, na.rm = TRUE),
          psychol = mean(psychol_21 / (popn_pums_21/100000)),
          high_income = sum(Household_Income >= 8, na.rm = TRUE),
          total_inc = sum(Household_Income == Household_Income, na.rm = TRUE), 
          prop_high_income = high_income/total_inc, 
          high_edu = sum(Educational_Attainment >= 6, na.rm = TRUE),
          total_edu = sum(Educational_Attainment == Educational_Attainment, na.rm = TRUE),
          prop_high_edu = high_edu/total_edu
          )

v2020 <- load_variables(2020, "pl", cache = TRUE)

census_data <- get_decennial(
  geography = "state",
  variables = c(
      "pop_total" = "P1_001N"
    ) ,
  year = 2020,
  geometry = TRUE
) 

census_data <- census_data |>
  mutate(region = tolower(NAME))

map_data <- left_join(census_data, joint_data, by = "region")
```

Here are all of the different factors and how they change across different states. The maps show the income and education appear to be correlated on the state level as many states with higher average incomes also have higher average levels of education. Additionally, areas with higher average incomes and levels of education appear to also have improved access to care. However, from the maps, there is no clear correlation with mental wellbeing. We will see if this correlations continue throughout our analysis. 

The coord_sf element was found using [Stack Overflow](https://stackoverflow.com/questions/76121455/select-adequate-x-axis-limits-to-center-an-sf-object-plotted-with-ggplot-and-geo )
``` {r}
#| echo: false
#| message: false
#| warning: false
#| results: 'hide'

ggplot(map_data) + 
 geom_sf(aes(fill = prop_high_income)) +
 coord_sf(xlim=c(-60, -170)) + 
 scale_fill_gradient(low = "red",high = "white") + 
 theme_void()

ggplot(map_data) + 
 geom_sf(aes(fill = prop_high_edu)) +
 coord_sf(xlim=c(-60, -170)) + 
 scale_fill_gradient(low = "blue", high = "white") + 
 theme_void()

ggplot(map_data) + 
 geom_sf(aes(fill = mean_total_mh)) +
 coord_sf(xlim=c(-60, -170)) + 
 scale_fill_gradient(low = "green", high = "white") +
 theme_void()

ggplot(map_data) + 
 geom_sf(aes(fill = psychol)) +
 coord_sf(xlim=c(-60, -170)) + 
 scale_fill_gradient(low = "purple", high = "white") +
 theme_void()


```
In order to get a more accurate relationships between income, education level and mental well being, we can now plot mental well being against each of these factors. 

Used [Stack Overflow] (https://stackoverflow.com/questions/1330989/rotating-and-spacing-axis-labels-in-ggplot2) to edit the lables.

```{r}
#| echo: false
#| message: false
#| warning: false
#| results: 'hide'

data_clean <- read_rds("dataset/data_clean.rds") |>
  mutate(total_mentalhealth = Anxious_Frequency + Worry_Frequency + 
         Little_Interest_Frequency + Depressed_Frequency)

income_sorted <- data_clean |>
  group_by(Household_Income) |>
  summarize(mean_total_mh = mean(total_mentalhealth, na.rm = TRUE)) |>
  na.omit(Household_Income) |>
  filter(Household_Income > 0)

ggplot(income_sorted) +
  geom_col(aes(x= factor(Household_Income), y = mean_total_mh, fill = "red")) +
  scale_x_discrete(
    labels=c("Less than $25,000", "$25,000 - $34,999", "$35,000 - $49,999",
             "$50,000 - $74,999","$75,000 - $99,999", "$100,000 - $149,999", "$150,000 - $199,999", "$200,000 and above" )
  ) + 
  coord_flip () +
  labs(
    x = "Income Level",
    y = "Average Mental Wellbeing",
    title = "Relationship Between Income Level and Mental Health"
  ) +
    theme (legend.position = "none")

#ggplot(income_sorted) +
#  geom_point(aes(x=Household_Income, y = mean_total_mh), alpha = 0.01) +
#  geom_smooth(aes(x=Household_Income, y = mean_total_mh), method = "lm", se = FALSE) +
#  theme (legend.position = "none") +
#  labs(
#    x = "Household Income",
#    y = "Average Mental Wellbeing",
#    title = "Relationship Between Income and Mental Health"
#  )

edu_sorted <- data_clean |>
  group_by(Educational_Attainment) |>
  summarize(mean_total_mh = mean(total_mentalhealth, na.rm = TRUE))

ggplot(edu_sorted) +
  geom_col(aes(x= factor(Educational_Attainment), y = mean_total_mh), fill = "blue") +
  scale_x_discrete(
    labels=c("Less than High School", "Some High School", "High School Degree",
             "Some College","Associate's Degree", "Bachelor's Degree", "Graduate Degree")
  ) + 
  coord_flip () +
  labs(
    x = "Educational Level",
    y = "Average Mental Wellbeing",
    title = "Relationship Between Education Level and Mental Health"
  ) +
    theme (legend.position = "none")

#ggplot(edu_sorted) +
#  geom_point(aes(x=Educational_Attainment, y = mean_total_mh), alpha = 0.01) +
#  geom_smooth(aes(x=Educational_Attainment, y = mean_total_mh), method = "lm", se = FALSE) +
#  theme (legend.position = "none") +
#  labs(
#    x = "Level of Education",
#    y = "Average Mental Wellbeing",
#    title = "Relationship Between Education Level and Mental Health"
#  )
```

Here are the relationships between access to mental health  and income and education level of each step. Following our analysis from the maps, there appears to be a  positive correlation between both income and education level and access to mental health. This means higher income and education levels in a state correspond with higher numbers of psychologists. 

However, this data is also limited as it only looks on the state level not a small scale such as by city or county which may have more variation in number of psychologists, income, and education level. 

```{r}
#| echo: false
#| message: false
#| warning: false
#| results: 'hide'


ggplot(joint_data) +
  geom_point(aes(x=prop_high_income, y = psychol), alpha = 0.01) +
  geom_smooth(aes(x=prop_high_income, y = psychol), method = "lm", se = FALSE) +
  theme (legend.position = "none") +
  labs(
    x = "Proportion of Families with Income over $200,000",
    y = "Psychologists Per 100K",
    title = "Relationship Between Income and Access to Mental Health Care"
  )


ggplot(joint_data) +
  geom_point(aes(x=prop_high_edu, y = psychol), alpha = 0.01) +
  geom_smooth(aes(x=prop_high_edu, y = psychol), method = "lm", se = FALSE) +
  theme (legend.position = "none") +
  labs(
    x = "Proportion of Individuals with a College Degree",
    y = "Psychologists Per 100K",
    title = "Relationship Between Income and Access to Mental Health Care"
  )

```

Finally, here is the relationship between average mental health in a state and number of psychologists in the state. There appears to be a negative relationship. This means that an increase in average mental wellbeing of a state correlates with a decrease in access to mental health care. 

```{r}
#| echo: false
#| message: false
#| warning: false
#| results: 'hide'
ggplot(joint_data) +
  geom_point(aes(x=mean_total_mh, y = psychol), alpha = 0.01) +
  geom_smooth(aes(x=mean_total_mh, y = psychol), method = "lm", se = FALSE) +
  theme (legend.position = "none") +
  labs(
    x = "Mental Wellbeing",
    y = "Psychologists Per 100K",
    title = "Relationship Between Mental Wellbeing and Access to Mental Health Care"
  )

```


This comes from the file `big_picture.Rmd`.

